\textsf{Are frontier models superhuman chemists?}

### 2024-03-19 src/tex/ms.tex
update affiliations
cleanup

\begin{figure}
    \centering
    \includegraphics{figures/question_diversity.pdf}
    \caption{\textbf{Principal component projection of embeddings of questions in the \chembench corpus.} To obtain this figure, we embedded question and answers using the BART model (using other embeddings, such as of OpenAI's ada model leads to qualitatively similar results). We then project the embeddings into a two-dimensional space using \gls{pca}. We color the points based on a classification into topics.}
    \label{fig:question_diversity}
    \script{plot_question_diversity.py}
\end{figure}

@@

In this spider chart, the worst score for every dimension is zero (no question answered completely correctly) and the best score is one (all questions answered correctly). 
Thus, a larger colored area indicates a better performance. 
One can observe that this performance varies widely across models and topics. 
While computational chemistry and biochemistry receive relatively high scores for many models, this is not the cases for topics such as chemical safety or analytical chemistry.
In the subfield of analytical chemistry the prediction of the number of peaks observable in a \gls{nmr} spectrum proved difficult for the models (e.g., \variable{output/subset_scores/is_number_nmr_peaks_gpt4.txt} percent correct answers for GPT-4) 
While this question is relatively easy (\variable{output/human_subset_scores/is_number_nmr_peaks.txt} percent correct) for trained humans given a drawing of the compounds, models that are only shown the \gls{smiles} string of a compound have to use this to reason about the symmetry of the compound. 
This is mirrored in the questions about the number of isomers, which is models struggle with more than humans (\variable{output/subset_scores/is_number_of_isomers_gpt4.txt} percent for GPT-4 vs. \variable{output/human_subset_scores/is_number_of_isomers.txt} percent for humans).
These findings also shin an interesting light on the value of textbook-inspired questions. A subset of the questions in the \chembench are based on textbooks targeted at undergraduate students. 
On those questions the models tend to perform better than on some our semi-automatically constructed tasks.

@@

\begin{figure}
    \centering
    \includegraphics{figures/confidence_vs_performance_overall.pdf}
    \caption{\textbf{Relationship between confidence of the model in the answer and correctness.}}
    \label{fig:confidence_vs_performance}
    \script{joint_analysis_confidence_performance.py}
\end{figure}

@@

\section*{Data and code availability}
The code and data for \chembench is available at \url{https://github.com/lamalab-org/chem-bench} and archived at \url{XXX}.
The code for the app for our human baseline study is available at \url{https://github.com/lamalab-org/chem-bench-app}. 
To ensure reproducibility, this manuscript was generated using the \href{https://show-your.work/en/latest/}{showyourwork!} framework.\cite{Luger2021}
The code to rebuild the paper (including code for all figures and numbers next to which there is a GitHub icon) can be found at \url{https://github.com/lamalab-org/chembench-paper}. 
To facilitate reproduction, some intermediate results of the analysis are cached at \url{http://dx.doi.org/10.5072/zenodo.34706}.

>>>>

\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{figures/question_diversity.pdf}
    \caption{\textbf{Principal component projection of embeddings of questions in the \chembench corpus.} To obtain this figure, we embedded question and answers using the BART model (using other embeddings, such as of OpenAI's ada model leads to qualitatively similar results). We then project the embeddings into a two-dimensional space using \gls{pca}. We color the points based on a classification into topics.}
    \label{fig:question_diversity}
    \script{plot_question_diversity.py}
\end{figure}

@@

In this spider chart, the worst score for every dimension is zero (no question answered completely correctly) and the best score is one (all questions answered correctly). 
Thus, a larger colored area indicates a better performance. 
One can observe that this performance varies widely across models and topics. 
While computational chemistry and biochemistry receive relatively high scores for many models, this is not the cases for topics such as chemical safety or analytical chemistry.
In the subfield of analytical chemistry the prediction of the number of peaks observable in a \gls{nmr} spectrum proved difficult for the models (e.g., \variable{output/subset_scores/is_number_nmr_peaks_gpt4.txt} percent correct answers for GPT-4) 
While this question is relatively easy (\variable{output/human_subset_scores/is_number_nmr_peaks.txt} percent correct) for trained humans given a drawing of the compounds, models that are only shown the \gls{smiles} string of a compound have to use this to reason about the symmetry of the compound. 
This is mirrored in the questions about the number of isomers, which is models struggle with more than humans (\variable{output/subset_scores/is_number_of_isomers_gpt4.txt} percent for GPT-4 vs. \variable{output/human_subset_scores/is_number_of_isomers.txt} percent for humans).
These findings also shine an interesting light on the value of textbook-inspired questions. A subset of the questions in the \chembench are based on textbooks targeted at undergraduate students. 
On those questions the models tend to perform better than on some our semi-automatically constructed tasks.

@@

\begin{figure}
    \centering
    \includegraphics{figures/confidence_vs_performance_overall.pdf}
    \caption{\textbf{Relationship between confidence of the model in the answer and correctness.} For this analysis, we used verbalized confidence estimates from the model. To obtain those estimates, we prompted the models to return a confidence score on an ordinal scale. We then plot the correctness of the answers (which can be either completely correct or completely incorrect) against the confidence score.
        The stripplots show the individual data points, the lines show the mean.}
    \label{fig:confidence_vs_performance}
    \script{joint_analysis_confidence_performance.py}
\end{figure}

@@

\section*{Data and code availability}
The code and data for \chembench is available at \url{https://github.com/lamalab-org/chem-bench} and archived at \url{XXX}.
The code for the app for our human baseline study is available at \url{https://github.com/lamalab-org/chem-bench-app}. 
To ensure reproducibility, this manuscript was generated using the \href{https://show-your.work/en/latest/}{\\showyourwork} framework.\cite{Luger2021}
The code to rebuild the paper (including code for all figures and numbers next to which there is a GitHub icon) can be found at \url{\GitHubURL}. 
To facilitate reproduction, some intermediate results of the analysis are cached at \url{http://dx.doi.org/10.5072/zenodo.34706}.
